

Scenario 1
Correct Door is door 1.
The contenstant correctly chooses door 1.
Monty can open eiter door 2 or door 3 to reveal a goat.
Let Suppose he opens door 2.
Monty will offer the contestant the chance to change his choice to door 3.


Scenario 2
Correct Door is door 1.
The contenstant chooses door 2.
Monty will open door 3 to reveal a goat. He cant open the correct door, and he cant open the door the contestant selected.
Monty will offer the contestant the chance to change his choice to door 1.



If the contestant picks the correct door - Monty Hall can choose either of the two wrong doors to reveal the goat.


However if the contestant picks a ``wrong" door -  Monty Hall has only one remaining door that he can open.


First - we must determine the doors that Monty can't Open (call this set `StayClosed`).

The remaining doors are the doors 

StayClosed = np.union1d(Correct,Choice)
CanOpen  = np.setdiff1d(AllDoors, StayClosed)


We can use very simple mathematical functions to implement an elegant way of selecting the door from the CanOpen Set. 
This set will have either one or two member elements, depending in the scenarios described above.

 - `len` : Determine the number of set members in CanOpen (either 1 or 2 )
 - `random` : Generate a random number between 0 and 1
 - `floor` : Rounds a number down to the preceding integer on the number line.


Scenario 1`
AllDoors = np.array([1,2,3])
CorrectDoor = 1
ChoiceDoor = 1
StayClosed = np.union1d (CorrectDoor,ChoiceDoor)
CanOpen = np.setdiff1d(AllDoors, StayClosed)
len(CanOpen)

Scenario 2
AllDoors = np.array([1,2,3])
CorrectDoor = 1
ChoiceDoor = 2
StayClosed = np.union1d (CorrectDoor,ChoiceDoor)
CanOpen = np.setdiff1d(AllDoors, StayClosed)
len(CanOpen)

len1 = 1
len1 * np.random.random(10)
All numbers between 0 and 1
floor(len1 * np.random.random(10) )
Values are all 0 

len1 = 2
len2 * np.random.random(10)
All numbers between 0 and 2
floor(len2 * np.random.random(10) )
Values are either 0 or 1

We can use the number generated as an index to select the door that Monty Hall will open to reveal a goat.

RandomIndex

WillOpen[RandomIndex]

