SciPy
Scipy.org
Docs
SciPy v0.17.1 Reference Guide
index
modules
modules
next
previous
Statistical functions (scipy.stats)

This module contains a large number of probability distributions as well as a growing library of statistical functions.

Each univariate distribution is an instance of a subclass of rv_continuous (rv_discrete for discrete distributions):

rv_continuous([momtype, a, b, xtol, ...])	A generic continuous random variable class meant for subclassing.
rv_discrete([a, b, name, badvalue, ...])	A generic discrete random variable class meant for subclassing.
Continuous distributions
alpha	An alpha continuous random variable.
anglit	An anglit continuous random variable.
arcsine	An arcsine continuous random variable.
beta	A beta continuous random variable.
betaprime	A beta prime continuous random variable.
bradford	A Bradford continuous random variable.
burr	A Burr (Type III) continuous random variable.
cauchy	A Cauchy continuous random variable.
chi	A chi continuous random variable.
chi2	A chi-squared continuous random variable.
cosine	A cosine continuous random variable.
dgamma	A double gamma continuous random variable.
dweibull	A double Weibull continuous random variable.
erlang	An Erlang continuous random variable.
expon	An exponential continuous random variable.
exponnorm	An exponentially modified Normal continuous random variable.
exponweib	An exponentiated Weibull continuous random variable.
exponpow	An exponential power continuous random variable.
f	An F continuous random variable.
fatiguelife	A fatigue-life (Birnbaum-Saunders) continuous random variable.
fisk	A Fisk continuous random variable.
foldcauchy	A folded Cauchy continuous random variable.
foldnorm	A folded normal continuous random variable.
frechet_r	A Frechet right (or Weibull minimum) continuous random variable.
frechet_l	A Frechet left (or Weibull maximum) continuous random variable.
genlogistic	A generalized logistic continuous random variable.
gennorm	A generalized normal continuous random variable.
genpareto	A generalized Pareto continuous random variable.
genexpon	A generalized exponential continuous random variable.
genextreme	A generalized extreme value continuous random variable.
gausshyper	A Gauss hypergeometric continuous random variable.
gamma	A gamma continuous random variable.
gengamma	A generalized gamma continuous random variable.
genhalflogistic	A generalized half-logistic continuous random variable.
gilbrat	A Gilbrat continuous random variable.
gompertz	A Gompertz (or truncated Gumbel) continuous random variable.
gumbel_r	A right-skewed Gumbel continuous random variable.
gumbel_l	A left-skewed Gumbel continuous random variable.
halfcauchy	A Half-Cauchy continuous random variable.
halflogistic	A half-logistic continuous random variable.
halfnorm	A half-normal continuous random variable.
halfgennorm	The upper half of a generalized normal continuous random variable.
hypsecant	A hyperbolic secant continuous random variable.
invgamma	An inverted gamma continuous random variable.
invgauss	An inverse Gaussian continuous random variable.
invweibull	An inverted Weibull continuous random variable.
johnsonsb	A Johnson SB continuous random variable.
johnsonsu	A Johnson SU continuous random variable.
ksone	General Kolmogorov-Smirnov one-sided test.
kstwobign	Kolmogorov-Smirnov two-sided test for large N.
laplace	A Laplace continuous random variable.
levy	A Levy continuous random variable.
levy_l	A left-skewed Levy continuous random variable.
levy_stable	A Levy-stable continuous random variable.
logistic	A logistic (or Sech-squared) continuous random variable.
loggamma	A log gamma continuous random variable.
loglaplace	A log-Laplace continuous random variable.
lognorm	A lognormal continuous random variable.
lomax	A Lomax (Pareto of the second kind) continuous random variable.
maxwell	A Maxwell continuous random variable.
mielke	A Mielke’s Beta-Kappa continuous random variable.
nakagami	A Nakagami continuous random variable.
ncx2	A non-central chi-squared continuous random variable.
ncf	A non-central F distribution continuous random variable.
nct	A non-central Student’s T continuous random variable.
norm	A normal continuous random variable.
pareto	A Pareto continuous random variable.
pearson3	A pearson type III continuous random variable.
powerlaw	A power-function continuous random variable.
powerlognorm	A power log-normal continuous random variable.
powernorm	A power normal continuous random variable.
rdist	An R-distributed continuous random variable.
reciprocal	A reciprocal continuous random variable.
rayleigh	A Rayleigh continuous random variable.
rice	A Rice continuous random variable.
recipinvgauss	A reciprocal inverse Gaussian continuous random variable.
semicircular	A semicircular continuous random variable.
t	A Student’s T continuous random variable.
triang	A triangular continuous random variable.
truncexpon	A truncated exponential continuous random variable.
truncnorm	A truncated normal continuous random variable.
tukeylambda	A Tukey-Lamdba continuous random variable.
uniform	A uniform continuous random variable.
vonmises	A Von Mises continuous random variable.
vonmises_line	A Von Mises continuous random variable.
wald	A Wald continuous random variable.
weibull_min	A Frechet right (or Weibull minimum) continuous random variable.
weibull_max	A Frechet left (or Weibull maximum) continuous random variable.
wrapcauchy	A wrapped Cauchy continuous random variable.
Multivariate distributions
multivariate_normal	A multivariate normal random variable.
matrix_normal	A matrix normal random variable.
dirichlet	A Dirichlet random variable.
wishart	A Wishart random variable.
invwishart	An inverse Wishart random variable.
Discrete distributions
bernoulli	A Bernoulli discrete random variable.
binom	A binomial discrete random variable.
boltzmann	A Boltzmann (Truncated Discrete Exponential) random variable.
dlaplace	A Laplacian discrete random variable.
geom	A geometric discrete random variable.
hypergeom	A hypergeometric discrete random variable.
logser	A Logarithmic (Log-Series, Series) discrete random variable.
nbinom	A negative binomial discrete random variable.
planck	A Planck discrete exponential random variable.
poisson	A Poisson discrete random variable.
randint	A uniform discrete random variable.
skellam	A Skellam discrete random variable.
zipf	A Zipf discrete random variable.
Statistical functions
Several of these functions have a similar version in scipy.stats.mstats which work for masked arrays.

describe(a[, axis, ddof, bias, nan_policy])	Computes several descriptive statistics of the passed array.
gmean(a[, axis, dtype])	Compute the geometric mean along the specified axis.
hmean(a[, axis, dtype])	Calculates the harmonic mean along the specified axis.
kurtosis(a[, axis, fisher, bias, nan_policy])	Computes the kurtosis (Fisher or Pearson) of a dataset.
kurtosistest(a[, axis, nan_policy])	Tests whether a dataset has normal kurtosis
mode(a[, axis, nan_policy])	Returns an array of the modal (most common) value in the passed array.
moment(a[, moment, axis, nan_policy])	Calculates the nth moment about the mean for a sample.
normaltest(a[, axis, nan_policy])	Tests whether a sample differs from a normal distribution.
skew(a[, axis, bias, nan_policy])	Computes the skewness of a data set.
skewtest(a[, axis, nan_policy])	Tests whether the skew is different from the normal distribution.
kstat(data[, n])	Return the nth k-statistic (1<=n<=4 so far).
kstatvar(data[, n])	Returns an unbiased estimator of the variance of the k-statistic.
tmean(a[, limits, inclusive, axis])	Compute the trimmed mean.
tvar(a[, limits, inclusive, axis, ddof])	Compute the trimmed variance
tmin(a[, lowerlimit, axis, inclusive, ...])	Compute the trimmed minimum
tmax(a[, upperlimit, axis, inclusive, ...])	Compute the trimmed maximum
tstd(a[, limits, inclusive, axis, ddof])	Compute the trimmed sample standard deviation
tsem(a[, limits, inclusive, axis, ddof])	Compute the trimmed standard error of the mean.
nanmean(*args, **kwds)	nanmean is deprecated!
nanstd(*args, **kwds)	nanstd is deprecated!
nanmedian(*args, **kwds)	nanmedian is deprecated!
variation(a[, axis, nan_policy])	Computes the coefficient of variation, the ratio of the biased standard deviation to the mean.
find_repeats(arr)	Find repeats and repeat counts.
trim_mean(a, proportiontocut[, axis])	Return mean of array after trimming distribution from both tails.
cumfreq(a[, numbins, defaultreallimits, weights])	Returns a cumulative frequency histogram, using the histogram function.
histogram2(*args, **kwds)	histogram2 is deprecated!
histogram(*args, **kwds)	histogram is deprecated!
itemfreq(a)	Returns a 2-D array of item frequencies.
percentileofscore(a, score[, kind])	The percentile rank of a score relative to a list of scores.
scoreatpercentile(a, per[, limit, ...])	Calculate the score at a given percentile of the input sequence.
relfreq(a[, numbins, defaultreallimits, weights])	Returns a relative frequency histogram, using the histogram function.
binned_statistic(x, values[, statistic, ...])	Compute a binned statistic for one or more sets of data.
binned_statistic_2d(x, y, values[, ...])	Compute a bidimensional binned statistic for one or more sets of data.
binned_statistic_dd(sample, values[, ...])	Compute a multidimensional binned statistic for a set of data.
obrientransform(*args)	Computes the O’Brien transform on input data (any number of arrays).
signaltonoise(*args, **kwds)	signaltonoise is deprecated!
bayes_mvs(data[, alpha])	Bayesian confidence intervals for the mean, var, and std.
mvsdist(data)	‘Frozen’ distributions for mean, variance, and standard deviation of data.
sem(a[, axis, ddof, nan_policy])	Calculates the standard error of the mean (or standard error of measurement) of the values in the input array.
zmap(scores, compare[, axis, ddof])	Calculates the relative z-scores.
zscore(a[, axis, ddof])	Calculates the z score of each value in the sample, relative to the sample mean and standard deviation.
sigmaclip(a[, low, high])	Iterative sigma-clipping of array elements.
threshold(*args, **kwds)	threshold is deprecated!
trimboth(a, proportiontocut[, axis])	Slices off a proportion of items from both ends of an array.
trim1(a, proportiontocut[, tail, axis])	Slices off a proportion from ONE end of the passed array distribution.
f_oneway(*args)	Performs a 1-way ANOVA.
pearsonr(x, y)	Calculates a Pearson correlation coefficient and the p-value for testing non-correlation.
spearmanr(a[, b, axis, nan_policy])	Calculates a Spearman rank-order correlation coefficient and the p-value to test for non-correlation.
pointbiserialr(x, y)	Calculates a point biserial correlation coefficient and its p-value.
kendalltau(x, y[, initial_lexsort, nan_policy])	Calculates Kendall’s tau, a correlation measure for ordinal data.
linregress(x[, y])	Calculate a linear least-squares regression for two sets of measurements.
theilslopes(y[, x, alpha])	Computes the Theil-Sen estimator for a set of points (x, y).
f_value(*args, **kwds)	f_value is deprecated!
ttest_1samp(a, popmean[, axis, nan_policy])	Calculates the T-test for the mean of ONE group of scores.
ttest_ind(a, b[, axis, equal_var, nan_policy])	Calculates the T-test for the means of TWO INDEPENDENT samples of scores.
ttest_ind_from_stats(mean1, std1, nobs1, ...)	T-test for means of two independent samples from descriptive statistics.
ttest_rel(a, b[, axis, nan_policy])	Calculates the T-test on TWO RELATED samples of scores, a and b.
kstest(rvs, cdf[, args, N, alternative, mode])	Perform the Kolmogorov-Smirnov test for goodness of fit.
chisquare(f_obs[, f_exp, ddof, axis])	Calculates a one-way chi square test.
power_divergence(f_obs[, f_exp, ddof, axis, ...])	Cressie-Read power divergence statistic and goodness of fit test.
ks_2samp(data1, data2)	Computes the Kolmogorov-Smirnov statistic on 2 samples.
mannwhitneyu(x, y[, use_continuity, alternative])	Computes the Mann-Whitney rank test on samples x and y.
tiecorrect(rankvals)	Tie correction factor for ties in the Mann-Whitney U and Kruskal-Wallis H tests.
rankdata(a[, method])	Assign ranks to data, dealing with ties appropriately.
ranksums(x, y)	Compute the Wilcoxon rank-sum statistic for two samples.
wilcoxon(x[, y, zero_method, correction])	Calculate the Wilcoxon signed-rank test.
kruskal(*args, **kwargs)	Compute the Kruskal-Wallis H-test for independent samples
friedmanchisquare(*args)	Computes the Friedman test for repeated measurements
combine_pvalues(pvalues[, method, weights])	Methods for combining the p-values of independent tests bearing upon the same hypothesis.
ss(*args, **kwds)	ss is deprecated!
square_of_sums(*args, **kwds)	square_of_sums is deprecated!
jarque_bera(x)	Perform the Jarque-Bera goodness of fit test on sample data.
ansari(x, y)	Perform the Ansari-Bradley test for equal scale parameters
bartlett(*args)	Perform Bartlett’s test for equal variances
levene(*args, **kwds)	Perform Levene test for equal variances.
shapiro(x[, a, reta])	Perform the Shapiro-Wilk test for normality.
anderson(x[, dist])	Anderson-Darling test for data coming from a particular distribution
anderson_ksamp(samples[, midrank])	The Anderson-Darling test for k-samples.
binom_test(x[, n, p, alternative])	Perform a test that the probability of success is p.
fligner(*args, **kwds)	Perform Fligner-Killeen test for equality of variance.
median_test(*args, **kwds)	Mood’s median test.
mood(x, y[, axis])	Perform Mood’s test for equal scale parameters.
boxcox(x[, lmbda, alpha])	Return a positive dataset transformed by a Box-Cox power transformation.
boxcox_normmax(x[, brack, method])	Compute optimal Box-Cox transform parameter for input data.
boxcox_llf(lmb, data)	The boxcox log-likelihood function.
entropy(pk[, qk, base])	Calculate the entropy of a distribution for given probability values.
chisqprob(*args, **kwds)	chisqprob is deprecated!
betai(*args, **kwds)	betai is deprecated!
Circular statistical functions
circmean(samples[, high, low, axis])	Compute the circular mean for samples in a range.
circvar(samples[, high, low, axis])	Compute the circular variance for samples assumed to be in a range
circstd(samples[, high, low, axis])	Compute the circular standard deviation for samples assumed to be in the range [low to high].
Contingency table functions
chi2_contingency(observed[, correction, lambda_])	Chi-square test of independence of variables in a contingency table.
contingency.expected_freq(observed)	Compute the expected frequencies from a contingency table.
contingency.margins(a)	Return a list of the marginal sums of the array a.
fisher_exact(table[, alternative])	Performs a Fisher exact test on a 2x2 contingency table.
Plot-tests
ppcc_max(x[, brack, dist])	Calculate the shape parameter that maximizes the PPCC
ppcc_plot(x, a, b[, dist, plot, N])	Calculate and optionally plot probability plot correlation coefficient.
probplot(x[, sparams, dist, fit, plot])	Calculate quantiles for a probability plot, and optionally show the plot.
boxcox_normplot(x, la, lb[, plot, N])	Compute parameters for a Box-Cox normality plot, optionally show it.
Masked statistics functions
Statistical functions for masked arrays (scipy.stats.mstats)
scipy.stats.mstats.argstoarray
scipy.stats.mstats.betai
scipy.stats.mstats.chisquare
scipy.stats.mstats.count_tied_groups
scipy.stats.mstats.describe
scipy.stats.mstats.f_oneway
scipy.stats.mstats.f_value_wilks_lambda
scipy.stats.mstats.find_repeats
scipy.stats.mstats.friedmanchisquare
scipy.stats.mstats.kendalltau
scipy.stats.mstats.kendalltau_seasonal
scipy.stats.mstats.kruskalwallis
scipy.stats.mstats.ks_twosamp
scipy.stats.mstats.kurtosis
scipy.stats.mstats.kurtosistest
scipy.stats.mstats.linregress
scipy.stats.mstats.mannwhitneyu
scipy.stats.mstats.plotting_positions
scipy.stats.mstats.mode
scipy.stats.mstats.moment
scipy.stats.mstats.mquantiles
scipy.stats.mstats.msign
scipy.stats.mstats.normaltest
scipy.stats.mstats.obrientransform
scipy.stats.mstats.pearsonr
scipy.stats.mstats.plotting_positions
scipy.stats.mstats.pointbiserialr
scipy.stats.mstats.rankdata
scipy.stats.mstats.scoreatpercentile
scipy.stats.mstats.sem
scipy.stats.mstats.signaltonoise
scipy.stats.mstats.skew
scipy.stats.mstats.skewtest
scipy.stats.mstats.spearmanr
scipy.stats.mstats.theilslopes
scipy.stats.mstats.threshold
scipy.stats.mstats.tmax
scipy.stats.mstats.tmean
scipy.stats.mstats.tmin
scipy.stats.mstats.trim
scipy.stats.mstats.trima
scipy.stats.mstats.trimboth
scipy.stats.mstats.trimmed_stde
scipy.stats.mstats.trimr
scipy.stats.mstats.trimtail
scipy.stats.mstats.tsem
scipy.stats.mstats.ttest_onesamp
scipy.stats.mstats.ttest_ind
scipy.stats.mstats.ttest_onesamp
scipy.stats.mstats.ttest_rel
scipy.stats.mstats.tvar
scipy.stats.mstats.variation
scipy.stats.mstats.winsorize
scipy.stats.mstats.zmap
scipy.stats.mstats.zscore
scipy.stats.mstats.compare_medians_ms
scipy.stats.mstats.gmean
scipy.stats.mstats.hdmedian
scipy.stats.mstats.hdquantiles
scipy.stats.mstats.hdquantiles_sd
scipy.stats.mstats.hmean
scipy.stats.mstats.idealfourths
scipy.stats.mstats.kruskal
scipy.stats.mstats.ks_2samp
scipy.stats.mstats.median_cihs
scipy.stats.mstats.meppf
scipy.stats.mstats.mjci
scipy.stats.mstats.mquantiles_cimj
scipy.stats.mstats.rsh
scipy.stats.mstats.sen_seasonal_slopes
scipy.stats.mstats.trimmed_mean
scipy.stats.mstats.trimmed_mean_ci
scipy.stats.mstats.trimmed_std
scipy.stats.mstats.trimmed_var
scipy.stats.mstats.ttest_1samp
Univariate and multivariate kernel density estimation (scipy.stats.kde)
gaussian_kde(dataset[, bw_method])	Representation of a kernel-density estimate using Gaussian kernels.
For many more stat related functions install the software R and the interface package rpy.

Table Of Contents
Statistical functions (scipy.stats)
Continuous distributions
Multivariate distributions
Discrete distributions
Statistical functions
Circular statistical functions
Contingency table functions
Plot-tests
Masked statistics functions
Univariate and multivariate kernel density estimation (scipy.stats.kde)
Previous topic
scipy.special.sinc

Next topic
scipy.stats.rv_continuous

© Copyright 2008-2014, The Scipy community.  Last updated on May 12, 2016.  Created using Sphinx 1.2.3.
